---
title: "DataGeneration.R"
author: "Meg Britt"
output: html_document
---
```{r}
library("knitr")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Create an object called Species that is a character vector containing the names of 10 species (you can use any species names you want)
```{r}
Species<-c("Pseudacris crucifer", "Homo sapiens","Ascaphus truei","Bombina bomina","Discoglossus galganoi","Megaphrys nastua","Pelobates fuscus","Hymenochirus boettgeri","Rhinophrynus dorsalis","Gastrotheca excubitor")
```

Create a vector called Limb.Width that is a numeric vector containing 10 random numbers sampled from a normal distribution with a mean of 10 and standard deviation of 10
```{r}
limb.width <- rnorm(10, mean=10, sd=10)
```

Convert to positive numbers
```{r}
limb.width<-abs(limb.width)
```

Create a vector called unitsW containing 10 strings that are either mm or cm generated randomly with a 0.5 probability of each
```{r}
unitsW <- sample(c("mm","cm"), 10, replace=TRUE, prob = c(0.5, 0.5))
```

Create a vector called Limb.Length that is a numeric vector containing 10 random numbers sampled from a normal distribution with a mean of 100 and standard deviation of 100 and convert to absolute values.
```{r}
limb.length <- rnorm(10, mean=100, sd=100)
limb.length<-abs(limb.length)
```

Create a vector called unitsL containing 10 strings that are either mm or cm generated randomly with a 0.5 probability of each
```{r}
unitsL <- sample(c("mm","cm"), 10, replace=TRUE, prob = c(0.5, 0.5))
```
Combine all vectors into a data.frame object called MyData
```{r}
MyData<-data.frame(Species,limb.width,unitsW,limb.length,unitsL, stringsAsFactors = FALSE)
```

Save MyData as a csv file called measurements.csv
```{r}
write.csv(MyData, "Data/measurements.csv")
```



